Debiased Calibration for Generalized Two-Phase Sampling

### 2025-01-20 final/paper/paper.tex
Updated paper with Dr. Kim corrections.

@@

The goal of statistical data integration is to combine information from one
survey with the variable of interest to another survey that only contains some
shared covariates. 
This integration usually involves one of two contexts: incorporating item
level responses or using summary statistics. These mirror the micro and macro
approaches described in \cite{yang2020statistical} respectively. The main
technique to incorporate item level responses is mass imputation. With access to
individual item level responses, the traditional approach is to build a
predictive model of the variable of interest from the covariates shared between
the two surveys and predict the missing outcome variable in the survey without
the variable of interest. An early use of mass imputation is found in
\cite{breidt1996two} where they used mass imputation to improve estimation in
the NRI survey. \cite{kim2012combining} used a parametric model to develop a
novel resampling technique to construct confidence intervals of population level
characteristics. \cite{chen2022nonparametric} use a nonparametric model to
estimate population characteristics. In the macro approach, one uses summary
statistics from other studies to adjust the current estimation. This is commonly
done using generalized least squares (GLS). 
\cite{hidiroglou2001double} applied the GLS technique to non-nested two-phase
samples. \cite{merkouris2004combining} used GLS with multiple surveys and showed
an optimality result. This was later extended to area level domains in 
\cite{merkouris2010combining}.

In this paper, we incorporate summary statistics of other surveys by
using a generalized two-phase sample similar to the GLS method. The first phase
sample consists of data
from multiple sources, and the second phase sample contains the existing
data. Unlike previous data integration methods, however, we also include a
pre-chosen function of the second phase design weights like
\cite{kwon2024debiased}. This allows us to develop theory showing that our
estimates are asymptotically design unbiased and that we have valid confidence
intervals. Moreover, we can show that this technique outperforms traditional
regression methods in finite samples.

This paper is organized as follows. Section 2 describes the basic setup of data
integration using summary statistics. Section 3 details the generalized
two-phase sampling framework and provides theory for our estimator being design
unbiased and the variance estimation to be valid. In Section 4, we extend this
framework first to non-nested two-phase sampling and second to the multiple
samples. Section 5 provides several simulation studies which show that our
method works, and Section 6 applies the method the Medicare Current Beneficiary
Survey.

@@

Consider a finite population, $U$, of size $N$ 
with measurement  $(\bx_i, y_i)$, where $\bx_i$ is the realization of the
auxiliary variables and $y_i$ is the realization of the study variable of
interest for element  $i$ in the finite population. The auxiliary variables are
cheap to measure but the measurement for the study variable is expensive.
Two-phase sampling is a popular method for cost effective sampling.
In the two-phase sampling, the first phase sample $A_1$  is selected from the
target population to obtain the measurement of $\bx_i$ in the sample. The
sampling design for the first-phase sample can be denoted by $P(A_1)$. After
that, based on the observed auxiliary information in the first-phase  sample,
the second-phase sample $A_2 \subseteq A_1$ is selected from the first-phase
sample to observe $y_i$. The sampling design for the second-phase sample is
denoted as $P( A_2 \mid A_1)$.  
The
goal of two-phase sampling is to construct an estimator of $Y_N =  \sum_{i=1}^N y_i$ 
that uses both the observed information in the Phase 2 sample and also the extra
auxiliary information from $\bf X$ in the Phase 1 sample. How to combine
information efficiently is an important problem in two-phase sampling.

Let $\pi_{1i} = \sum_{A_1: i \in A_1} \Pr(A_1)$ be the first-order inclusion
probability for the first-phase sample and let $\pi_{2i | 1} = \sum_{A_2: i \in
A_2} \Pr(A_2 \mid A_1)$ be the first-order inclusion probability for the
second-phase sample. From the first-phase sample, we can compute
$\widehat{\mathbf{X}}_1 = \sum_{i \in A_1} \pi_{1i}^{-1} \bx_i$ as an unbiased
estimator of $\mathbf{X}_N = \sum_{i=1}^N  \bx_i$. Also, from the second-phase
sample, we can compute $\left( \widehat{\mathbf{X}}_2 , \hat{Y}_2 \right) =
\sum_{i \in A_2} \pi_{1i}^{-1} \pi_{2i \mid 1}^{-1} \left( \bx_i, y_i\right)$ as
an unbiased estimator of $(\mathbf{X}_N, Y_N)$.
We therefore have three estimators for
two parameters. The two-phase regression estimator can be  used to combine this
information efficiently and it can be expressed as  
\begin{equation} \label{eq:1}
\hat Y_{\rm reg, tp} 
= \sum_{i \in A_1} \frac{1}{\pi_{1i}} \bf x_i^\top  \hat{\bm \beta_q} + 
\sum_{i \in A_2} \frac{1}{\pi_{1i}\pi_{2i|1}} \left(y_i - \bf x_i^\top  \hat{\bm \beta_q}\right)
\end{equation} 
where $q_i = q(\bf x_i)$  is a function of $\bf x_i$ and  
$$
\hat{\bm \beta_q} = \left(\sum_{i \in A_2} 
 \frac{1}{\pi_{1i} q_i } \bf x_i \bf x_i^\top \right)^{-1} 
\sum_{i \in A_2} \frac{1}{\pi_{1i} q_i} \bf x_i y_i .
$$ 
Regarding the choice of $q_i$, one can consider a linear regression model 
\begin{equation}
y_i = \bx_i^\top \bm \beta + e_i 
\end{equation}
where $e_i \sim \left(0, \sigma^2 c_i \right)$, where $c_i=c( \bx_i)$ is a known
function. In this case,   $q_i=c_i$ is a reasonable choice.

One way to understand the regression estimator is to use the following decomposition:

$$ 
\hat Y_{\rm reg, tp} 
= \underbrace{\sum_{i \in A_1} \frac{\bf x_i \hat{\bm \beta_q}}{\pi_{1i}}}_{
  \text{ Prediction  }} + \underbrace{\sum_{i \in A_2}
\frac{1}{\pi_{1i}\pi_{2i|1}} (y_i - \bf x_i \hat{\bm \beta_q})}_{
\text{Bias correction}}.
$$
The prediction term is obtained by finding the best linear unbiased predictor
under the working superpopulation model. The bias correction term 
makes the two-phase regression estimator asymptotically design unbiased
regardless of the underlying the superpopulation model. 
Note that the two-phase regression estimator in (\ref{eq:1})  can be
expressed as  $\hat Y_{\rm reg, tp} =
\sum_{i \in A_2} \hat w_{2i} y_i / \pi_{1i}$ where

@@

This means that $\hat Y_{reg, tp}$ is also a calibration estimator. The idea
that regression estimation is a form of calibration was noted by
\cite{deville1992calibration} and extended by them to consider loss functions
other than just squared loss. Their generalized loss function minimizes
$\sum_i G(w_i, d_i)q_i$ for weights $w_i$ and design-weights $d_i$ where
$G(\cdot)$ is a non-negative, strictly convex function with respect to $w$,
defined on an interval containing $d_i$, with $g(w_i, d_i) = \partial G /
\partial w$ continuous.\footnote{The \cite{deville1992calibration} paper
considers regression estimators for a single phase setup, which we apply to our
two-phase example.} This
generalization includes empirical likelihood estimation, and maximum entropy
estimation among others. The variance estimation is based on a linearization
that shows that minimizing the generalized loss function subject to the
calibration constraints is asymptotically equivalent to a regression estimator.

@@

To construct asymptotically unbiased estimate of the variance we assume that the
sampling designs for $A_1$ and $A_2$ are measurable and we define the following
joint inclusion probabilities:

$\Pr(i \in A_1, j \in A_1) = \pi_{1ij}, \text{ and }
\Pr(i \in A_2, j \in A_2 \mid i \in A_1, j \in A_1) = \pi_{2ij|1}.$

@@

We follow the approach of \cite{kwon2024debiased} for the debiased calibration
method.  Let $G: \mathcal{V} \to \R$ be the generalized entropy function that is
strictly convex and differentiable. We consider maximizing

\begin{equation}\label{eq:primalloss}
  H(w) = - \sum_{i \in A_2} \frac{1}{\pi_{1i}} G(w_{2i}) q_i
\end{equation}

subject to the following constraints:

\begin{equation}\label{eq:calconst1}
  \sum_{i \in A_2} \frac{\bf x_i w_{2i}}{\pi_{1i}} = 
\sum_{i \in A_1} \frac{\bf x_i}{\pi_{1i}}
\end{equation}

and

\begin{equation}\label{eq:calconst2}
  \sum_{i \in A_2} \frac{g(\pi_{2i|1}^{-1})w_{2i}q_i}{\pi_{1i}} = 
  \sum_{i \in A_1} \frac{g(\pi_{2i|1}^{-1})q_i}{\pi_{1i}}, 
\end{equation}
where $g(w) = \partial G / \partial w$.

The first constraint is the existing calibration constraint and the second
ensures that design consistency is achieved. 
The original method of \cite{kwon2024debiased} only considers known finite
population quantities on the right hand side of \eqref{eq:calconst1}.

By writing $w_{1i} = \pi_{1i}^{-1}$, the the goal is to solve

\begin{equation}\label{eq:primal}
  \hat w_2  = \argmin_{w_{2}} \sum_{i \in A_2} w_{1i} G(w_{2i}) q_i 
  \text{ such that}
  \sum_{i \in A_2} w_{1i} w_{2i} \bf z_i q_i = \sum_{i \in A_1} w_{1i} \bf z_i q_i
\end{equation}

where  $\bf z_i = (\bf x_i / q_i, g(\pi_{2i|1}^{-1}))$. The resulting 
estimator of $Y_N$ is

\begin{equation}\label{eq:tpest}
\hat Y_{\rm DCE} = \sum_{i \in A_2} w_{1i} \hat w_{2i} y_i.
\end{equation}

@@

We can understand the estimated weights, $\hat w_{2i}$ from
Equation~\ref{eq:primal} using the method of Lagrange multipliers. We need to
minimize the Lagrangian function

\begin{equation}\label{eq:legragedc1}
  L(w_{2i}, \bm \lambda) = \sum_{i \in A_2} w_{1i} G(w_{2i}) q_i 
  + \bm \lambda \left( \sum_{i \in A_1} w_{1i} \bf z_i q_i -
    \sum_{i \in A_2} w_{1i} w_{2i} \bf z_i q_i\right)
\end{equation}

where $\bm \lambda$ is a vector of Lagrange multipliers.
Differentiating with respect to $w_{2i}$ and setting this expression equal to
zero, yields the fact that $\hat w_{2i}$ satisfies

$$ \hat w_{2i} = g^{-1}(\hat{\bm \lambda}^T \bf z_i) $$

where $\hat{\bm \lambda}$ is the solution to

\begin{equation}\label{eq:lamdc1}
  \left( \sum_{i \in A_1} w_{1i} \bf z_i q_i -
  \sum_{i \in A_2} w_{1i} w_{2i}(\hat{\bm \lambda}) \bf z_i q_i\right) = 0.
\end{equation}

The first result we show is that the construction of the weights in
Equation~\ref{eq:primal} leads to a design consistent estimator 
$\hat Y_{\rm DCE}$ in Equation~\ref{eq:tpest}.

\begin{theorem}[Design Consistency]\label{thm:dc1}
  Let $\bm \lambda^*$ be the probability limit of $\hat{\bm \lambda}$. 
  Under some regularity conditions,

  $$
  \hat Y_{\rm DCE} = \hat Y_\ell(\hat{\bm{\lambda}}, \bm \phi^*) + O_p(N / n_2)
  $$

  where

  $$
  \hat Y_{\ell}(\hat{\bm \lambda}, \bm \phi^*) = 
  \sum_{i \in A_2} w_{1i} \hat w_{2i} y_i
  + \left(\sum_{i \in A_1} w_{1i} \bf z_i q_i - 
    \sum_{i \in A_2} w_{1i} \hat w_{2i} \bf{z}_i q_i\right)\bm \phi^*,
  $$

  and

  $$
  \bm \phi^* = 
  \left[\sum_{i \in U} \frac{\pi_{2i|1}\bf z_i \bf z_i^T q_i}{g'(d_{2i|1})}\right]^{-1}
  \sum_{i \in U} \frac{\pi_{2i|1}\bf z_i y_i}{g'(d_{2i|1})}.
  $$

%   \begin{align*}
%   \hat Y_{\ell}(\bm \lambda^*, \bm \phi^*) 
%   &= \hat{Y}_{\rm DCE} + \left(\sum_{i \in A_1} w_{1i} \bf x_i  -
%     \sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1} \bf x_i  \right)^T \bm \phi_1^* \\
%   &\quad+ \left(\sum_{i \in A_1} w_{1i} g_i  -
%  \sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1}g_i  \right)^T \phi_2^*  
%  \end{align*} 
%\textcolor{red}{(You introduced $\hat{Y}_{\ell}$ twice!)}
%  and  
%
%  $$
%  \begin{pmatrix}
%  \bm \phi_1^* \\
%  \phi_2^* 
%  \end{pmatrix}
%  = \left[ \sum_{i \in U} \frac{\pi_{2i \mid 1} }{ g'(d_{2i|1}) q_i} 
%  \begin{pmatrix}
%  \bx_i \bx_i^T &   \bx_i g_i   \\
%  g_i  \bx_i^T   & g_i^2     
%  \end{pmatrix} \right]^{-1}
%  \sum_{i \in U} \frac{\pi_{2i|1}}{ g'(d_{2i | 1}) q_i} 
%  \begin{pmatrix}
%    \bx_i \\ g_i 
%  \end{pmatrix}y_i 
%  $$

@@

See the appendix for the proof of Theorem 1. Notice that this theorem implies
that $\hat Y_{\ell}(\hat{\bm \lambda}, \bm \phi^*)$ and $\hat Y_{\rm DCE}$ have
the same asymptotic expectation and variance, which means that we can estimate
the variance of $\hat Y_{\ell}(\hat{\bm \lambda}, \bm \phi^*)$ to get an
estimate of the variance of $\hat Y_{\rm DCE}$. To estimate $\bm \phi^*$, we can
use

$$
\hat{\bm \phi} = 
\left[\sum_{i \in A_2} 
w_{1i}\frac{\bf z_i \bf z_i^T q_i}{g'(d_{2i|1})}\right]^{-1}
\sum_{i \in A_2} w_{1i}\frac{\bf z_i y_i}{g'(d_{2i|1})}.
$$

The next theorem discusses how to compute an estimate for the variance of 
$\hat Y_{\rm DCE}$. For the proof see the appendix.

@@

  Define $\hat \eta_i = \bf z_i q_i \hat{\bm \phi} + 
  \frac{\delta_i}{\pi_{2i|1}}(y_i - \bf z_i q_i \hat{\bm \phi})$.
  We can construct an unbiased estimate the variance of $\hat Y_{\rm DCE}$ with

  $$
  \hat V_{\rm DCE} 
  = \sum_{i \in A_1} \sum_{j \in A_1} \frac{\Delta_{1ij}}{\pi_{1ij}} 
  \hat \eta_i \hat \eta_j
  + \sum_{i \in A_2} \sum_{j \in A_2} w_{1i} \frac{\Delta_{2ij|1}}{\pi_{2ij|1}} 
  \frac{(y_i - \bf z_i q_i \hat{\bm \phi})}{\pi_{2i|1}}
  \frac{(y_j - \bf z_j q_j \hat{\bm \phi})}{\pi_{2j|1}}
  $$

  where $\Delta_{1ij} = \pi_{1ij} - \pi_{1i}\pi_{1j}$ and 
  $\Delta_{2ij|1} = \pi_{2ij|1} - \pi_{2i|1}\pi_{2j|1}$.

@@

Consider the setup in which we have independent samples $A_0, A_1, \dots, A_M$
where $Y$ is observed only in $A_0$ but $\bf X$ is observed in all of the
samples with variables $\bf X^{(0)}$ observed in $A_0$ and variables $\bf X^{(m)}$
observed in $A_m$ for each $m = 1, \dots, M$. Like the non-nested case, we
assume that each survey is sampled independently from the same population
sampling frame.

>>>>

Updates:
\begin{itemize}
  \item Modified the end of paragraph two of the introduction and the beginning
    of paragraph three of the introduction to mirror more of the content of the
    rest of the paper.
  \item Fixed spacing issues with the beginning of paragraphs.
  \item Added Table~\ref{tab:msexample}
\end{itemize}
\newpage

@@

The goal of statistical data integration is to combine information from one
survey with the variable of interest to another survey with
shared covariates. 
This integration usually involves one of two contexts: incorporating item
level responses or using summary statistics. These mirror the micro and macro
approaches described in \cite{yang2020statistical} respectively. The main
technique to incorporate item level responses is mass imputation. With access to
individual item level responses, the traditional approach is to build a
predictive model of the variable of interest from the covariates shared between
the two surveys and predict the missing outcome variable in the survey without
the variable of interest. An early use of mass imputation is found in
\cite{breidt1996two} where they use mass imputation to improve estimation in
the National Resources Inventory survey. Later, \cite{kim2012combining} use a
parametric model to develop a novel resampling technique to construct confidence
intervals of population level characteristics. \cite{chen2022nonparametric} use
a nonparametric model to estimate population characteristics. In the macro
approach, one uses summary statistics from other studies to adjust the current
estimation. This is commonly done using generalized least squares (GLS).
\cite{hidiroglou2001double} apply the GLS technique to non-nested two-phase
samples. \cite{merkouris2004combining} use GLS with multiple surveys and showed
an optimality result. This was later extended to area level domains in
\cite{merkouris2010combining}. The main idea of the GLS approach is to construct
a regression estimator for two-phase samples. In Section 2 of the paper, we
explain the basic approach of regression estimation and show how it can be
generalized to a calibration estimator.

This paper, however, generalizes the two-phase regression estimator using the
debiased calibration technique of \cite{kwon2024debiased}.
We incorporate summary statistics of other surveys by
using a generalized two-phase sample similar to the GLS method. The first phase
sample consists of data
from multiple sources, and the second phase sample contains the existing
data. Unlike previous data integration methods, however, we also include a
pre-chosen function of the second phase design weights, which allows us to
develop theory showing that our
estimates are asymptotically design unbiased and that we have valid confidence
intervals. Moreover, we can show that this technique outperforms traditional
regression methods in finite samples.

\textcolor{red}{(I am not sure whether the current introduction provides a good
  guide for the main contents of the paper. The introduction does not match
  well with the main contents of the paper.)
}

\textcolor{red}{CL: I have modified the last two paragraphs to help tie into the
main work of the paper.}

This paper is organized as follows. 
Section 2 describes the existing methods to combine probability samples using
regression estimators or calibration.
Section 3 details the generalized
two-phase sampling framework and provides theory for our estimator to be design
unbiased and the variance estimation to be valid. In Section 4, we extend this
framework first to non-nested two-phase sampling and second to multiple
samples. Section 5 provides several simulation studies which show that our
method works, and Section 6 applies the method to the Medicare Current
Beneficiary Survey.

@@

Consider a finite population, $U = \{1, \dots, N\}$, of size $N$ 
with measurement  $(\bx_i, y_i)$, where $\bx_i$ is the realization of the
auxiliary variables and $y_i$ is the realization of the study variable of
interest for element  $i$ in the finite population. The auxiliary variables are
cheap to measure but the measurement for the study variable is expensive.
Two-phase sampling is a popular method for cost effective sampling.
In two-phase sampling, the first phase sample $A_1$  is selected from the
target population to obtain the measurement of $\bx_i$ in the sample. The
sampling design for the first-phase sample can be denoted by $P(A_1)$. After
that, based on the observed auxiliary information in the first-phase  sample,
the second-phase sample $A_2 \subseteq A_1$ is selected from the first-phase
sample to observe $y_i$. The sampling design for the second-phase sample is
denoted as $P( A_2 \mid A_1)$.  
The
goal of two-phase sampling is to construct an estimator of $Y_N =  \sum_{i=1}^N y_i$ 
that uses both the observed information in the Phase 2 sample and also the 
auxiliary information from $\bf X \in \R^p$ in the Phase 1 sample. How to
combine information efficiently is an important problem in two-phase sampling.

Let $\pi_{1i} = \sum_{A_1: i \in A_1} \Pr(A_1)$ be the first-order inclusion
probability for the first-phase sample and let $\pi_{2i | 1} = \sum_{A_2: i \in
A_2} \Pr(A_2 \mid A_1)$ be the first-order inclusion probability for the
second-phase sample. From the first-phase sample, we can compute
$\widehat{\mathbf{X}}_1 = \sum_{i \in A_1} \pi_{1i}^{-1} \bx_i$ as an unbiased
estimator of $\mathbf{X}_N = \sum_{i=1}^N  \bx_i$. Also, from the second-phase
sample, we can compute $\left( \widehat{\mathbf{X}}_2 , \hat{Y}_2 \right) =
\sum_{i \in A_2} \pi_{1i}^{-1} \pi_{2i \mid 1}^{-1} \left( \bx_i, y_i\right)$ as
an unbiased estimator of $(\mathbf{X}_N, Y_N)$.
We therefore have three estimators for
two parameters. The two-phase regression estimator can be  used to combine this
information efficiently and it can be expressed as  
\begin{equation} \label{eq:1}
\hat Y_{\rm reg, tp} 
= \sum_{i \in A_1} \frac{1}{\pi_{1i}} \bf x_i^\top  \hat{\bm \beta_q} + 
\sum_{i \in A_2} \frac{1}{\pi_{1i}\pi_{2i|1}} \left(y_i - \bf x_i^\top  \hat{\bm \beta_q}\right)
\end{equation} 
where $q_i = q(\bf x_i)$  is a function of $\bf x_i$ and

$$
\hat{\bm \beta_q} = \left(\sum_{i \in A_2} 
 \frac{1}{\pi_{1i} q_i } \bf x_i \bf x_i^\top \right)^{-1} 
\sum_{i \in A_2} \frac{1}{\pi_{1i} q_i} \bf x_i y_i .
$$

Regarding the choice of $q_i$, one can consider a linear regression model 
\begin{equation}
y_i = \bx_i^\top \bm \beta + e_i 
\label{regression}
\end{equation}
where $e_i \sim \left(0, \sigma^2 c_i \right)$, where $c_i=c( \bx_i)$ is a known
function. In this case,  $q_i = c_i$ is a reasonable choice. 
One way to understand the two-phase regression estimator is to use the following
decomposition:

$$ 
\hat Y_{\rm reg, tp} 
= \underbrace{\sum_{i \in A_1} \frac{\bf x_i^\top \hat{\bm \beta_q}}{\pi_{1i}}}_{
  \text{ Prediction  }} + \underbrace{\sum_{i \in A_2}
\frac{1}{\pi_{1i}\pi_{2i|1}} (y_i - \bf x_i^\top \hat{\bm \beta_q})}_{
\text{Bias correction}}.
$$
The prediction term is obtained by finding the best linear unbiased predictor
under the working superpopulation model. The bias correction term 
makes the two-phase regression estimator asymptotically design unbiased
regardless of the underlying the superpopulation model. 
Note that the two-phase regression estimator in Equation~\ref{eq:1} can be
expressed as  $\hat Y_{\rm reg, tp} =
\sum_{i \in A_2} \pi_{1i}^{-1} \hat w_{2i} y_i$ where

@@

This means that $\hat Y_{\rm reg, tp}$ is also a calibration estimator. The idea
that regression estimation is a form of calibration was noted by
\cite{deville1992calibration} and extended by them to consider loss functions
other than squared loss. Their generalized loss function minimizes
$\sum_i G(w_i, d_i)q_i$ for weights $w_i$ and design-weights $d_i$ where
$G(\cdot)$ is a non-negative, strictly convex function with respect to $w$,
defined on an interval containing $d_i$, with $g(w_i, d_i) = \partial G /
\partial w$ being continuous.\footnote{The \cite{deville1992calibration} paper
considers regression estimators for a single phase setup, which we apply to our
two-phase example.} This
generalization includes empirical likelihood estimation, and maximum entropy
estimation among others \citep{schennach2007point}.

@@

@@

We follow the approach of \cite{kwon2024debiased} for the debiased calibration
method.  Let $G: \mathcal{V} \to \R$ be the generalized entropy function that is
strictly convex and differentiable. We consider maximizing 
\begin{equation}\label{eq:primalloss}
  H(w_2) = - \sum_{i \in A_2} \frac{1}{\pi_{1i}} G(w_{2i}) q_i
\end{equation}
subject to the following constraints:
\begin{equation}\label{eq:calconst1}
  \sum_{i \in A_2} \frac{\bf x_i w_{2i}}{\pi_{1i}} = 
\sum_{i \in A_1} \frac{\bf x_i}{\pi_{1i}}
\end{equation}
and 
\begin{equation}\label{eq:calconst2}
  \sum_{i \in A_2} \frac{g(\pi_{2i|1}^{-1})w_{2i}q_i}{\pi_{1i}} = 
  \sum_{i \in A_1} \frac{g(\pi_{2i|1}^{-1})q_i}{\pi_{1i}}, 
\end{equation}
where $g(w) = \partial G / \partial w$.

The first constraint is the usual calibration constraint motivated from the
regression model in (\ref{regression}) and the second
ensures that design consistency is achieved. 
The original method of \cite{kwon2024debiased} only considers known finite
population quantities on the right hand side of \eqref{eq:calconst1}.
By writing $w_{1i} = \pi_{1i}^{-1}$, the the goal is to solve
\begin{equation}\label{eq:primal}
  \hat w_2  = \argmin_{w_{2}} \sum_{i \in A_2} w_{1i} G(w_{2i}) q_i 
  \text{ such that}
  \sum_{i \in A_2} w_{1i} w_{2i} \bf z_i q_i = \sum_{i \in A_1} w_{1i} \bf z_i q_i
\end{equation}
where  $\bf z_i = (\bf x_i / q_i, g(\pi_{2i|1}^{-1}))$. The resulting 
estimator of $Y_N$ is 
\begin{equation}\label{eq:tpest}
\hat Y_{\rm DCE} = \sum_{i \in A_2} w_{1i} \hat w_{2i} y_i.
\end{equation}

@@

We can understand the estimated weights, $\hat w_{2i}$ from
Equation~\ref{eq:primal} using the method of Lagrange multipliers. We need to
minimize the Lagrangian function
\begin{equation}\label{eq:legragedc1}
  L(w_{2i}, \bm \lambda) = \sum_{i \in A_2} w_{1i} G(w_{2i}) q_i 
  + \bm \lambda^\top \left( \sum_{i \in A_1} w_{1i} \bf z_i q_i -
    \sum_{i \in A_2} w_{1i} w_{2i} \bf z_i q_i\right)
\end{equation}
where $\bm \lambda$ is a vector of Lagrange multipliers.
Differentiating with respect to $w_{2i}$ and setting this expression equal to
zero, we can obtain 
$$ \hat w_{2i} = g^{-1}(\hat{\bm \lambda}^\top \bf z_i) $$
where $g(w) = \partial G / \partial w$ and 
$\hat{\bm \lambda}$ is the solution to
\begin{equation}\label{eq:lamdc1}
  \left( \sum_{i \in A_1} w_{1i} \bf z_i q_i -
  \sum_{i \in A_2} w_{1i} w_{2i}(\hat{\bm \lambda}) \bf z_i q_i\right) = 0.
\end{equation}
The following theorem shows that the construction of the weights in
Equation~\ref{eq:primal} leads to a design consistent estimator 
$\hat Y_{\rm DCE}$ in Equation~\ref{eq:tpest}.

\begin{theorem}[Design Consistency]\label{thm:dc1}
  Let $\bm \lambda^*$ be the probability limit of $\hat{\bm \lambda}$. 
  Under the regularity conditions stated in the appendix
  $$
  \hat Y_{\rm DCE} = \hat Y_\ell(\bm{\lambda}^*, \bm \phi^*) + O_p(N / n_2)
  $$
  where
  $$
  \hat Y_{\ell}(\bm \lambda^*, \bm \phi^*) = 
  \sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1} y_i
  + \left(\sum_{i \in A_1} w_{1i} \bf z_i q_i - 
  \sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1} \bf{z}_i q_i\right)\bm \phi^*,
  $$
  and
  $$
  \bm \phi^* = 
  \left[\sum_{i \in U} \frac{\pi_{2i|1}\bf z_i \bf z_i^\top q_i}{g'(d_{2i|1})}\right]^{-1}
  \sum_{i \in U} \frac{\pi_{2i|1}\bf z_i y_i}{g'(d_{2i|1})}.
  $$

%   \begin{align*}
%   \hat Y_{\ell}(\bm \lambda^*, \bm \phi^*) 
%   &= \hat{Y}_{\rm DCE} + \left(\sum_{i \in A_1} w_{1i} \bf x_i  -
%     \sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1} \bf x_i  \right)^\top \bm \phi_1^* \\
%   &\quad+ \left(\sum_{i \in A_1} w_{1i} g_i  -
%  \sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1}g_i  \right)^\top \phi_2^*  
%  \end{align*} 
%\textcolor{red}{(You introduced $\hat{Y}_{\ell}$ twice!)}
%  and  
%
%  $$
%  \begin{pmatrix}
%  \bm \phi_1^* \\
%  \phi_2^* 
%  \end{pmatrix}
%  = \left[ \sum_{i \in U} \frac{\pi_{2i \mid 1} }{ g'(d_{2i|1}) q_i} 
%  \begin{pmatrix}
%  \bx_i \bx_i^\top &   \bx_i g_i   \\
%  g_i  \bx_i^\top   & g_i^2     
%  \end{pmatrix} \right]^{-1}
%  \sum_{i \in U} \frac{\pi_{2i|1}}{ g'(d_{2i | 1}) q_i} 
%  \begin{pmatrix}
%    \bx_i \\ g_i 
%  \end{pmatrix}y_i 
%  $$

@@

See the appendix for the proof of Theorem~\ref{thm:dc1}. Notice that this
theorem implies
that $\hat Y_{\ell}(\bm \lambda^*, \bm \phi^*)$ and $\hat Y_{\rm DCE}$ have
the same asymptotic expectation and variance, 
which means that we can estimate
the variance of $\hat Y_{\ell}(\bm \lambda^*, \bm \phi^*)$ to get an
estimate of the variance of $\hat Y_{\rm DCE}$. The basic idea of 
Theorem~\ref{thm:dc1} is that asymptotically the two-phase debiased calibration
method is equivalent to a regression estimator. By constraints 
(\ref{eq:calconst1}) and (\ref{eq:calconst2}), 
$\hat Y_{\ell}(\bm \lambda^*, \bm \phi^*)$ is design consistent, and the
variance is asymptotically equivalent to the two-phase regression estimator. 
We can estimate $\bm \phi^*$ with
\begin{equation}
\hat{\bm \phi} = 
\left[\sum_{i \in A_2} 
w_{1i}\frac{\bf z_i \bf z_i^\top q_i}{g'(d_{2i|1})}\right]^{-1}
\sum_{i \in A_2} w_{1i}\frac{\bf z_i y_i}{g'(d_{2i|1})}.
\label{eq:8}
\end{equation}
Therefore, the proposed calibration estimator in (\ref{eq:primal}) is
asymptotically equivalent to the two-phase regression estimator given by 
  $$
  \hat Y_{\rm tp, reg}  = 
  \sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1} y_i
  + \left(\sum_{i \in A_1} w_{1i} \bf z_i q_i - 
  \sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1} \bf{z}_i q_i\right)\hat{\bm \phi},
  $$
where $\hat{\bm \phi}$ is defined in (\ref{eq:8}).

For specific sample designs, we can construct the optimal entropy function
$G(\cdot)$. From Theorem~\ref{thm:dc1}, the choice of entropy function affects
the variance of the debiased calibration estimator through the form of $g'(d_{2i|1})$
in $\bm \phi^*$. Consider a class of generalized regression estimators,
$$
\mathbf{B} = 
\left\{\sum_{i \in A_1} w_{1i} \bf{z}_i^\top \bm \gamma + 
\sum_{i \in A_2} \frac{w_{1i}}{\pi_{2i|1}}(y_i - \bf{z}_i^\top \bm \gamma):
\bm \gamma \in \R^{p+1}\right\}
$$
Then the design optimal estimator is 
$$
\hat \theta_{\ell, \rm{opt}} = 
\sum_{i \in A_1} w_{1i} \bf{z}_i^\top \bm \phi_{\rm opt} + 
\frac{w_{1i}}{\pi_{2i|1}}\sum_{i \in A_2} (y_i - \bf{z}_i^\top \bm \phi_{\rm opt}):
$$
where 
$$
\bm \phi_{\rm opt} = 
\Var\left(\sum_{i \in A_2} d_{2i|1} \bf{z}_i\right)^{-1} 
\Cov\left(\sum_{i \in A_2} d_{2i|1} \bf{z}_i, \sum_{i \in A_2}, d_{2i|1} y_i \right).
$$
We assume that both $A_1$ and $A_2$ are Poisson samples. Then,
$$
\bm \phi_{\rm opt} = 
\left\{
\sum_{i \in U} \left(d_{2i|1} - 1 + d_{2i|1}(d_{1i} - 1)\right) \bf{z_i}\bf{z_i}^\top
\right\}^{-1}
\sum_{i \in U} \left(d_{2i|1} - 1 + d_{2i|1}(d_{1i} - 1)\right) \bf{z_i} y_i.
$$
To find the optimal entropy function, we solve the differential equation
$$
\frac{1}{d_{2i|1} g'(d_{2i|1})} = d_{2i|1} - 1 + d_{2i|1}(d_{1i} - 1).
$$
Using the fact that $(y(xy-1))^{-1} = -y^{-1} + x(xy-1)^{-1}$, the optimal
entropy function is 
$$
G(w_{2i|1}) = - \log(w_{2i|1}) + \log(w_{2i|1}w_{1i} - 1).
$$
To construct an asymptotically unbiased estimate of the variance we assume that
the sampling designs for $A_1$ and $A_2$ are measurable and we define the
following joint inclusion probabilities:
$$
\Pr(i \in A_1, j \in A_1) = \pi_{1ij}, \text{ and }
\Pr(i \in A_2, j \in A_2 \mid i \in A_1, j \in A_1) = \pi_{2ij|1}.
$$
The next theorem discusses how to compute an estimate for the variance of 
$\hat Y_{\rm DCE}$. For the proof see the appendix.

@@

  Define $\hat \eta_i = \bf z_i q_i \hat{\bm \phi} + 
  \frac{\delta_i}{\pi_{2i|1}}(y_i - \bf z_i q_i \hat{\bm \phi})$.
  We can construct an unbiased estimate the variance of $\hat Y_{\rm DCE}$ with
  $$
  \hat V_{\rm DCE} 
  = \sum_{i \in A_1} \sum_{j \in A_1} \frac{\Delta_{1ij}}{\pi_{1ij}} 
  \hat \eta_i \hat \eta_j
  + \sum_{i \in A_2} \sum_{j \in A_2} w_{1i} \frac{\Delta_{2ij|1}}{\pi_{2ij|1}} 
  \frac{(y_i - \bf z_i q_i \hat{\bm \phi})}{\pi_{2i|1}}
  \frac{(y_j - \bf z_j q_j \hat{\bm \phi})}{\pi_{2j|1}}
  $$
  where $\Delta_{1ij} = \pi_{1ij} - \pi_{1i}\pi_{1j}$ and 
  $\Delta_{2ij|1} = \pi_{2ij|1} - \pi_{2i|1}\pi_{2j|1}$.
\end{theorem}

@@

Consider the setup in which we have independent samples $A_0, A_1, \dots, A_M$
where $Y$ is observed only in $A_0$ but $\bf X$ is observed in all of the
samples with variables $\bf X^{(0)}$ observed in $A_0$ and variables $\bf X^{(m)}$
observed in $A_m$ for each $m = 1, \dots, M$. Like the non-nested case, we
assume that each survey is sampled independently from the same population
sampling frame. Some basic examples of data structures that would benefit from
the multi-source two-phase sampling can be found in Table~\ref{tab:msexample}.
